---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- load libraries -->
```{r , include = FALSE}
library(tidyverse)
library(TidyWrappers)
```

# TidyWrappers 

<!-- badges: start -->
```{r , echo=FALSE , results='asis' , message=FALSE}
cat(
badger::badge_devel("cparsania/tidywrappers" , color = "blue"),
badger::badge_lifecycle()
)

```

<!-- badges: end -->

TidyWrappers is an R package to deal with an object `tbl`. Utilities given here are appropriate to use when a `tbl` has rows are features / observations and columns are variables. It provides handy wrapper functions on top of `dplyr` verbs to get, count, convert, keep, remove and replace data from a `tbl` object.  


## Motivation

The way R is being used, has changed drastically in last few years due to the availability of [`tidyverse`](https://www.tidyverse.org/packages/){target="_blank"} and more specifically [`dplyr`](https://dplyr.tidyverse.org){target="_blank"} (all credits to [Hadley Wickham](http://hadley.nz) and [RStudio](https://rstudio.com) team). Rich documentation, stable release and excellent functionalities of these packages have provoked significant number of R users to switch from traditional `dataframes` to intelligible data structure, `tibble` (aka `tbl`). `dplyr` provides users with rich utilities like `select`, `mutate`, `filter`, `group_by` , `summarise` and `arrange` for easy manipulation of a `tibble`.  However, for newbie it takes a while to become familiar with `tidy data` and `dplyr` core and helper functions.  

`TidyWrappers` contains set of functions, which save you little from writing and implementing  core `dplyr` verbs while manipulating data from `tbl`. Additionaly, redundant lines of code can be avoided using functions given in `TidyWrappers`.  There are more than 30 functions implemented in `TidyWrappers`, which wrap various `dplyr` functions internally. See examples below.


```{r}
  tbl <- tibble::tibble(a = letters[1:6],x = c(0,1,2,0,3,5) , y = c(0,1,2,0,3,5) , z = c(0,1,2,0,3,5) )
  tbl
  
  ## keep rows having  0 across all numeric columns. 
  
  # using dplyr
  
  tbl %>% dplyr::filter_if(is.numeric , dplyr::any_vars( . == 0) )
  
  # using TidyWrappers
  
  tbl %>% tbl_keep_rows_zero_any()

  ## remove rows having  0 across all numeric columns. 
  
  # using dplyr 
  
  tbl %>% dplyr::filter_if(is.numeric , dplyr::any_vars( . > 0) )
  
  # using TidyWrappers 
  
  tbl %>% TidyWrappers::tbl_remove_rows_zero_any()
  
```

## Install

Install current version of TidyWrappers on your system.

```{r, eval=FALSE}
# install.packages("devtools")
library(devtools)
devtools::install_github("cparsania/TidyWrappers", dependencies = TRUE)

```


## Functions 

### Convert tbl operations  - `tbl_convert_*`

`tbl_convert_*` funtions return a `tbl`. Functions apply to only numeric columns of a input `tbl`.

```{r , eval = FALSE}
tbl_convert_column_zscore()	## Scale numeric columns. 
tbl_convert_log()	## Apply log on numeric columns.
tbl_convert_log10()	## Apply log10 on numeric columns.
tbl_convert_log2()	## Apply log2 on numeric columns.
tbl_convert_row_zscore()	## Scale rows considering numeric columns.
```
### Count columns operations - `tbl_count_vars*`

`tbl_count_*` functions return a numeric value. Functions apply to both numeric and non numeric columns of a `tbl`.

```{r, eval = FALSE}
tbl_count_vars_NA_all()	## Count columns having all values are NA.
tbl_count_vars_NA_any()	## Count columns having atleast one NA.
tbl_count_vars_zero_all()	## Count columns having all values are 0.
tbl_count_vars_zero_any()	## Count columns having atleast one 0.
```

### Get vars operations - `tbl_get_vars*`

`tbl_get_vars*` functions returns a character vector of column names. Functions apply to both numeric and non numeric columns of a `tbl`.

```{r, eval  = FALSE}
tbl_get_vars_NA_all()	## Get columns names having all values are NA.
tbl_get_vars_NA_any()	## Get columns names having atleast one NA.
tbl_get_vars_zero_all()	## Get columns names having all values are 0.
tbl_get_vars_zero_any()	## Get columns names having atleast one 0.
```
### Keep rows operations - `tbl_keep_rows*`

`tbl_keep_rows*` functions returns a `tbl`. Functions apply to numeric columns of a `tbl`.

```{r, eval = FALSE}
tbl_keep_rows_less_than_or_equal_all()	## Keep rows with less than or equal 'cutoff' in all numeric columns.
tbl_keep_rows_less_than_or_equal_any()	## Keep rows with less than or equal 'cutoff' in any numeric column.
tbl_keep_rows_greater_than_or_equal_all()	## Keep rows with greater than or equal 'cutoff' in all numeric columns.
tbl_keep_rows_greater_than_or_equal_any()	## Keep rows with greater than or equal 'cutoff' in any numeric column.
tbl_keep_rows_zero_all()	## Keep rows having all values are 0.
tbl_keep_rows_zero_any()	## Keep rows having atleast one 0.
```

### Remove rows operations - `tbl_remove_rows*`

`tbl_remove_rows*` functions returns a `tbl`. Functions apply to numeric columns of a `tbl`.

```{r , eval = FALSE}
tbl_remove_rows_greater_than_or_equal_any()	## Remove rows with greater than or equal 'cutoff' in any numeric column.
tbl_remove_rows_less_than_or_equal_any()	## Remove rows with less than or equal 'cutoff' in any numeric column.
tbl_remove_rows_greater_than_or_equal_all()	## Remove rows with greater than or equal 'cutoff' in all numeric column.
tbl_remove_rows_less_than_or_equal_all()	## Remove rows with less than or equal 'cutoff' in all numeric column.
tbl_remove_rows_zero_all()	## Remove rows having all values are 0.
tbl_remove_rows_zero_any()	## Remove rows having atleast one 0
```

### Remove columns operations - `tbl_remove_vars*`

`tbl_remove_vars*` functions returns a `tbl`. Functions apply to numeric and non columns of a `tbl`.

```{r, eval=FALSE}
tbl_remove_vars_NA_all()	## Remove columns having all values are NA.
tbl_remove_vars_NA_any()	## Remove columns having atlease one NA
tbl_remove_vars_zero_all()	## Remove columns having all values are 0.
tbl_remove_vars_zero_any()	## Remove columns having atleast one 0
```

### Replace value operations - `tbl_replace*`

`tbl_replace*` functions returns a `tbl`. Functions apply to both numeric and non numeric columns of a `tbl`.

``` {r , eval = FALSE}
tbl_replace_greater_than()	## Replace numeric values greater than given 'cutoff'.
tbl_replace_greater_than_or_equal()	## Replace numeric values greater than or equal to given 'cutoff'.
tbl_replace_less_than()	## Replace numeric values less than given 'cutoff'.
tbl_replace_less_than_or_equal()	## Replace numeric values less than or equal given 'cutoff'.
tbl_replace_string()	## Replace all occurance of 'pattern' in all non numeric columns. Wrapper around 'stringr::str_replace_all()'.


```


